---
title: "Take-Home-Exercise_04"
---

1. Data Rangling

```{r}
#Unpivot the data
pacman::p_load(plotly, DT, patchwork, ggstatsplot, tidyverse)
library(tidyr)
library(dplyr)
library(readr)
```

```{r}
imports <- read_csv("data/imports.csv")
exports <- read_csv("data/exports.csv")
```

```{r}
imports_recent <- imports %>%
  select(matches("Data Series|2022\\s\\w+|2021\\s\\w+|2020\\s\\w+"))
imports_recent
```

```{r}
imports_transpose <- imports_recent %>% 
  pivot_longer(
    cols = !`Data Series`,
    names_to = "Months", 
    values_to = "Imports")

head(imports_transpose)
```

```{r}
exports_recent <- exports %>%
  select(matches("Data Series|2022\\s\\w+|2021\\s\\w+|2020\\s\\w+"))

exports_transpose <- exports_recent %>% 
  pivot_longer(
    cols = !`Data Series`,
    names_to = "Months", 
    values_to = "Exports"
    )

head(exports_transpose)
```
```{r}
export_line <- ggplot(exports_transpose, aes(x = Months, y = Exports, color = "Data Series")) +
  geom_line() +
  labs(x = "Month", y = "Amount")

export_line
```


```{r}

# merge imports and exports datasets by date
data_all <- merge(imports_recent, exports_recent, by = "Data Series")
head(data_all)
```

```{r}
data_all_transpose <- data_all %>% 
  pivot_longer(
    cols = !`Data Series`,
    names_to = "Months", 
    values_to = "Amount"
    )
head(data_all_transpose)
```

```{r}
# plot line chart of imports and exports
ggplot(data_all, aes(x = "Months", color = c("\\s\\w+.x", "\\s\\w+.y"))) +
  geom_line() +
  labs(x = "Month", y = "Amount") +
  scale_color_manual(values = c("red", "blue"), labels = c("Imports", "Exports")) +
  theme_bw()
```




For the purpose of this take-home exercise, Merchandise Trade provided by Department of Statistics, Singapore (DOS) will be used. The data are available under the sub-section of Merchandise Trade by Region/Market. You should download the data by clicking on the link Download all in Excel on the same webpage. The study period should be between January 2020 to December 2022.

The Designing Tool
For the purpose of this take-home exercise, ggplot2 and its extension should be used to design the analytical visualisation. tidyverse family of packages should be used to prepare the data.

The Write-up
The write-up of the take-home exercise should include but not limited to the followings:

Describe the selection and designed consideration of the analytical data visualisation used. The discussion should limit to not more than 150 words each.

A reproducible description of the procedures used to prepare the analytical visualisation. Please refer to the peer submission I shared.

A write-up of not more than 100 words to discuss the patterns reveal by each analytical visualisation prepared